---
title: "Rearrangement signatures regression "
author: "Alvin"
date: "08/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(caret)
```

## Regression analysis to associate rearrangement signatures (RS) to molecular features

This document contains the code to run a logistic regression using the presence of the RS (RS*.bin) in each tumour and features include : 
- RS Signature weights, log-scaled 
- Number of Me events, log-scaled
- SNV.subtypes from Secrier 2016
- Number of complex SV clusters, log-scaled
- Number of chromothripsis events
- Chemotherapy status of specimen
- ecDNA amplicons affecting driver genes, 1/0
- Number of mutations assigned to SNV signatures, log-scaled 
- Number of breakage-fusion-bridge cycle events, log-scaled
- Number of ecDNA events, log-scaled

```{r, functions}
## function for logistic regression
log.reg <- function(response, reg) {
  bin.res <- paste0(response, '.bin')
  univariate <- lapply(colnames(reg)[!grepl(paste('Id', response, bin.res, '.bin', 'RS.', sep = '|'), colnames(reg))], function(x){
    
    suppressMessages(
      fit <- glm( as.formula(paste0(paste(bin.res, '~', sep=' '), x)), data = reg, family = 'binomial', maxit = 1000)
    )
    coef <- summary(fit)$coefficient
    
    ## univariate fdr adjusted p < 0.05
    if ((coef[2,4] < 0.05) ){ 
      coef <- coef[2,, drop=F]
      coef
    } else {
      coef <-  NULL }
  }
  )
  
  uni<- do.call(rbind,compact(univariate))
  uni<- cbind(uni, p.adjust(uni[,4], 'fdr'))
  
  if (nrow(uni) > 0){ 
    
    ## multivariate fdr adjusted p < 0.05
    coef <- rownames(uni)[ uni[,5] < 0.05]
    coef <- coef[!is.na(coef)]
  } else coef <- c()
  
  if (length(coef) > 0 ){ 
    form <- as.formula(paste( c(paste(bin.res, '~', sep=' '), gsub('DDR|Mutagenic', '', coef)), collapse = '+'))
    out.model <- glm(form, data = reg, family = 'binomial', maxit = 1000)
    out.model <- stepAIC(out.model, reg )
  }
}

## function to create partitions for each RS, 10 replicates each
part.list <- function(x, percent, reg) {
  
  index <- createDataPartition(reg[,x], p = percent,times =  10)
  
  lapply(index, function(y){
    
    train.set <- reg %>% dplyr::slice(y)
    test.set <- reg %>% dplyr::filter(!Id %in% train.set$Id)
    
    list(train = train.set, 
         test = test.set )
  })
}

## generate logistic regression model
generate.model <- function(sv, partition) {
  lapply(partition, function(y){
    log.reg(sv, y[['train']])
  })
}


## m is model 
generate.tables <- function(m){ 
  
  resample.names <- names(m)
  
  out <- bind_rows(lapply(resample.names, function(n){ 
    
    df <- m[[n]]
    
    if (!is.null(df)){ 
      
      output <- cbind(  as.data.frame(coefficients(summary(df)))[,1:4],
                        exp(coef(df)), exp(confint(df)))
      
      output %>%  rownames_to_column('Variable') %>% mutate(index = n )
    }
  }))
  out %>% filter(`Pr(>|z|)` < 0.05 & Variable !='(Intercept)')
}


```

## Regression model run on all samples in the paper 
```{r, full.model}
reg.data <- readRDS('rds/regression.input.rds')
## SV signatures to run
svs <- colnames(reg.data)[grepl('RS.*\\d$', colnames(reg.data))]

## Run the regression 
final.model <- lapply(svs, function(x) log.reg(x, reg.data))

## Generate final summary tablmes
final.tables <- lapply(final.model, function(x){ 
  
  if (!is.null(x)){ 

  output <- cbind(  as.data.frame(coefficients(summary(x)))[,1:4],
                     exp(coef(x)), exp(confint(x)))

    output
    
    }
})

names(final.tables) <- svs

## output table and filter by fdr-adjusted p-values
comb.table <- do.call(rbind, final.tables) %>% rownames_to_column('Variable') %>% 
  arrange(desc(`exp(coef(x))`)) %>%  
  mutate(padj = p.adjust(`Pr(>|z|)`, 'fdr')) %>% 
  filter(padj < 0.1 & !grepl('Intercept', Variable))

## Supp table output 
comb.table %>% filter(`exp(coef(x))` > 1.5 | `exp(coef(x))` < 0.5 )
```

## Regression using RNAseq expression of driver genes 
Expression data from RNAs is input as TPM without scaling 
```{r, expression}
exp.reg <- readRDS('rds/regression.expression.input.rds')
exp.model <- lapply(svs, function(x) log.reg(x, exp.reg))

exp.tables <- lapply(exp.model, function(x){ 
  
  if (!is.null(x)){ 

  output <- cbind(  as.data.frame(coefficients(summary(x)))[,1:4],
                     exp(coef(x)), exp(confint(x)))

    output
    
    }
})

names(exp.tables) <- svs
exp.comb.tables <- do.call(rbind, exp.tables) %>% rownames_to_column('Variable') %>%  
  filter(!grepl('Intercept', Variable)) %>%
  mutate(padj = p.adjust(`Pr(>|z|)`, 'fdr')) %>%  
  filter(padj < 0.1 ) %>% 
  arrange(desc(`exp(coef(x))`))

## expression output in supp table
exp.comb.tables %>% filter(`exp(coef(x))` > 1.5 | `exp(coef(x))`< 0.5)
```

## Generate subsets of cohort to determine robustness of associated features
Generate partitions of 80-20, 60-40, 40-60 and test how many times each feature is significant in 10 replicates per partition. 

```{r, partitions}
## generate sample partitions for 80, 60, 40 
parts <- c(1, .8, .6, .4)
partitions <- lapply(setNames(parts, nm = parts), function(p) {
  partition <- lapply(svs, function(s) part.list(s, p, reg.data %>% dplyr::select(-Chemo_status)))
  names(partition) <- svs
  partition
    }
  ) 

## run regression for each partition, replicate and summarize freq per variable 
out <- lapply(setNames(names(partitions), nm = names(partitions)), function(x){
  sv.names <- names(partitions[[x]])
  output <- lapply(setNames(sv.names, nm = sv.names), function(sv){ 
     generate.tables(generate.model(sv, partitions[[x]][[sv]])) %>% 
      group_by(Variable) %>% 
      dplyr::summarise("counts.{x}" := n()/10) 
    
    })
  
  output
  
})

## summarize output and from each partition and compare output per signature
out.summary <- lapply(setNames(svs, nm = svs), function(x) {
  df  <- list(out[['1']][[x]], out[['0.8']][[x]], out[['0.6']][[x]], out[['0.4']][[x]]) %>% 
    reduce(full_join, by = "Variable")
   df %>% arrange(counts.1) 
  }
)

## output, 1 df per RS
out.summary

```









